### 2025-02-21 products/manuscript/Manuscript.qmd

---
title: "Factors affecting health sector expenditure in OECD countries"
author: "Prasanga Paudel"
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

This uses HTML as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like docx or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(here)
library(readxl)
library(writexl)
```

**Authors**

* Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678) 
* Second Author$^{3, *}$
* Third Autor$^{3, \dagger}$
* Last Author$^{1, 2, \land}$

**Author affiliations**

1. College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.
2. Another Affiliation, Athens, GA, USA.
3. Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: some@email.com

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract
_The data is based on World Development Indicator from World Bank Group. This is a comprehensive database that includes information including employment, public and private expenditure in health sector, fertility, immunization, literacy, mortality, and many more.The dataset has a total of 1496 variables and across 11 years from 2010 to 2020. The actual source of data has information from 1960 to 2023 but to deal with the problem of missing values in recent years and overcomplication in analysis due to too many year data , we will focus on 10 recent years between 2010 and 2020._

{{< pagebreak >}}

# Introduction

## General Background Information
_We will try to analyze the most influential factors affecting the expenditure in health sector by different countries. The predictor variables will include factors ranging from education, mortality, population distribution among rural and urban spaces, and so on. We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors for this panel data._

_We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure. Some research have also focused on the relation between life expectancy and health expenditure. However, these research were based on fixed effect and random effect models. In our analysis we will try to follow OLS multiple regression model._

## Description of data and data source
_This is a time series data based on 38 member countries from the Organisation for Economic Co-operation and Development (OECD). This organization is a unique forum with governments from 37 democracies collaborate to develop policy standards to promote sustainable economic growth. However, the data obtained from this dataset is crude and needs to be treated before attempting analysis based on it. For example, in its present form, the variable name is a long sentense describing the data it contains.The years are listed as column names and variables are listed as Series._

## Objective of the research

The expenditure in health sector is one of the important factor that helps in determining the efficiency of health sector. We can relate the expenditure with different national development indicators and hope to find important connections between them. This will help government to make specific interventions through their policy to enhance their health sector. We can also find out how countries are currently operating and which factor has helped or negatively affected their health sector.

_The main objective of this research is to:_

_1. To analyze the factors that affect the expenditure in health sector among OECD countries._

_2. To analyze the magnitude of the influence of these factors on health expenditure_

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

## Data aquisition
_The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime. We can apply various filters as per our need and export the data as a .CSV or an .xlsx file._

Link to the source: <https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#>

# Methods

_To be added_

# Data import and cleaning
```{r}
file_path <- here("data","raw-data", "wdi-oecd.xlsx")
data <- read_excel(file_path, sheet="Data")
```

## A look into the original for of dataset

_This is the snapshot of the data as of now._
```{r}
head(data)
```

# EDA based on clean dataset

## Descriptive Analysis

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

### Looking into the trend for our possible dependent variable

@fig-health-expenditure-plot how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.

```{r}
#| label: fig-health-expenditure-plot
#| fig-cap: "Current Health Expenditure as a % of GDP (2010–2020): Top 5 nations"
#| echo: FALSE
knitr::include_graphics(here("results","figures","health_expenditure_plot.png"))
```

@fig-health-exp-aus how "Health Expenditure as a Percentage of GDP changes over time for Australia.

```{r}
#| label: fig-health-exp-aus
#| fig-cap: "Health Expenditure (% of GDP), Country: Australia"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "health_exp_aus.png"))
```

@fig-pop-growth-aus checks if Health Expenditure as a Percentage of GDP over the Years shows any reation with annual population growth for Australia.

```{r}
#| label: fig-pop-growth-aus
#| fig-cap: "Health Expenditure (% of GDP), Country: Australia"
#| echo: FALSE
knitr::include_graphics(here("results","figures","pop_growth_aus.png"))
```

@fig-he_per_capita checks if Health Expenditure Per Capita (USD) over the Years shows any reation with annual population growth for Australia.

```{r}
#| label: fig-he_per_capita
#| fig-cap: "Health Expenditure Per Capita (USD, Country: Australia"
#| echo: FALSE
knitr::include_graphics(here("results","figures","health_exp_per_capita_aus.png"))
```

## Basic statistical analysis

_To be added_

## Full analysis

_To be added._

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_To be added._

## Strengths and Limitations
_To be added._

## Conclusions
_To be added._

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References