### 2025-03-12 code/analysis-code/statistical-analysis.qmd
project part 3

---
title: "statistical-analysis"
format: html
editor: visual
---

## Quarto

```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

```{r}
# Load necessary libraries
library(dplyr)

# Define the dependent variable
dependent_var <- "Current_health_expenditure_per_capita_current_US"

# Select 50 variables that are likely to predict the dependent variable
selected_vars <- c(
  # Economic Factors
  "GDP_per_capita_current_US", "GNI_per_capita_Atlas_method_current_US", "GDP_growth_annual_",
  "Inflation_GDP_deflator_annual_", "Domestic_credit_provided_by_financial_sector__of_GDP",
  "Gross_capital_formation__of_GDP", "Exports_of_goods_and_services__of_GDP",
  "Imports_of_goods_and_services__of_GDP", "Foreign_direct_investment_net_inflows_BoP_current_US",
  "Industry_including_construction_value_added__of_GDP", "Agriculture_forestry_and_fishing_value_added__of_GDP",
  "Services_value_added__of_GDP", "Unemployment_total__of_total_labor_force_modeled_ILO_estimate",
  "Income_share_held_by_lowest_20", "Income_share_held_by_highest_20", "Poverty_headcount_ratio_at_215_a_day_2017_PPP__of_population",

  # Health Factors
  "Life_expectancy_at_birth_total_years", "Mortality_rate_under5_per_1000_live_births",
  "Prevalence_of_HIV_total__of_population_ages_1549", "Immunization_measles__of_children_ages_1223_months",
  "Births_attended_by_skilled_health_staff__of_total", "Contraceptive_prevalence_any_method__of_married_women_ages_1549",
  "Hospital_beds_per_1000_people", "Physicians_per_1000_people", "Prevalence_of_underweight_weight_for_age__of_children_under_5",
  "Adolescent_fertility_rate_births_per_1000_women_ages_1519", "Current_health_expenditure__of_GDP",
  "Domestic_general_government_health_expenditure__of_GDP",

  # Social Factors
  "Population_total", "Population_growth_annual_", "Population_density_people_per_sq_km_of_land_area",
  "Urban_population__of_total_population", "Rural_population__of_total_population",
  "Literacy_rate_adult_total__of_people_ages_15_and_above", "School_enrollment_primary__gross",
  "School_enrollment_secondary__gross", "School_enrollment_tertiary__gross", "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate",
  "Net_migration", "Refugee_population_by_country_or_territory_of_asylum",

  # Government Factors
  "Military_expenditure__of_GDP", "Tax_revenue__of_GDP", "Government_expenditure_on_education_total__of_GDP",
  "Government_Effectiveness_Estimate", "Rule_of_Law_Estimate", "Control_of_Corruption_Estimate",
  "Political_Stability_and_Absence_of_ViolenceTerrorism_Estimate", "Regulatory_Quality_Estimate",

  # Trade Factors
  "Trade__of_GDP", "Merchandise_trade__of_GDP", "Exports_of_goods_and_services_current_US",
  "Imports_of_goods_and_services_current_US", "Foreign_direct_investment_net_inflows_BoP_current_US",
  "External_debt_stocks_total_DOD_current_US", "Net_barter_terms_of_trade_index_2015__100"
)

# Add the dependent variable to the selected variables
selected_vars <- c(dependent_var, selected_vars)

# Subset the dataset with the selected variables
mydata_selected <- mydata %>%
  select(all_of(selected_vars))

# Categorize the selected variables
variable_categories <- list(
  Economic = c("GDP_per_capita_current_US", "GNI_per_capita_Atlas_method_current_US", "GDP_growth_annual_",
               "Inflation_GDP_deflator_annual_", "Domestic_credit_provided_by_financial_sector__of_GDP",
               "Gross_capital_formation__of_GDP", "Exports_of_goods_and_services__of_GDP",
               "Imports_of_goods_and_services__of_GDP", "Foreign_direct_investment_net_inflows_BoP_current_US",
               "Industry_including_construction_value_added__of_GDP", "Agriculture_forestry_and_fishing_value_added__of_GDP",
               "Services_value_added__of_GDP", "Unemployment_total__of_total_labor_force_modeled_ILO_estimate",
               "Income_share_held_by_lowest_20", "Income_share_held_by_highest_20", "Poverty_headcount_ratio_at_215_a_day_2017_PPP__of_population"),

  Health = c("Life_expectancy_at_birth_total_years", "Mortality_rate_under5_per_1000_live_births",
             "Prevalence_of_HIV_total__of_population_ages_1549", "Immunization_measles__of_children_ages_1223_months",
             "Births_attended_by_skilled_health_staff__of_total", "Contraceptive_prevalence_any_method__of_married_women_ages_1549",
             "Hospital_beds_per_1000_people", "Physicians_per_1000_people", "Prevalence_of_underweight_weight_for_age__of_children_under_5",
             "Adolescent_fertility_rate_births_per_1000_women_ages_1519", "Current_health_expenditure__of_GDP",
             "Domestic_general_government_health_expenditure__of_GDP"),

  Social = c("Population_total", "Population_growth_annual_", "Population_density_people_per_sq_km_of_land_area",
             "Urban_population__of_total_population", "Rural_population__of_total_population",
             "Literacy_rate_adult_total__of_people_ages_15_and_above", "School_enrollment_primary__gross",
             "School_enrollment_secondary__gross", "School_enrollment_tertiary__gross", "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate",
             "Net_migration", "Refugee_population_by_country_or_territory_of_asylum"),

  Government = c("Military_expenditure__of_GDP", "Tax_revenue__of_GDP", "Government_expenditure_on_education_total__of_GDP",
                 "Government_Effectiveness_Estimate", "Rule_of_Law_Estimate", "Control_of_Corruption_Estimate",
                 "Political_Stability_and_Absence_of_ViolenceTerrorism_Estimate", "Regulatory_Quality_Estimate"),

  Trade = c("Trade__of_GDP", "Merchandise_trade__of_GDP", "Exports_of_goods_and_services_current_US",
            "Imports_of_goods_and_services_current_US", "Foreign_direct_investment_net_inflows_BoP_current_US",
            "External_debt_stocks_total_DOD_current_US", "Net_barter_terms_of_trade_index_2015__100")
)

# Print the categorized variables
print(variable_categories)
```

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      # For tidy regression results
library(purrr)      # For functional programming
library(kableExtra) # For advanced table formatting

# Ensure all selected variables are numeric
mydata <- mydata %>%
  mutate(across(everything(), as.numeric))

# Function to run multiple regression for a specific category and country
run_regression <- function(data, category_vars, category_name) {
  # Drop variables with all NA values within the category
  category_vars <- category_vars[sapply(data[category_vars], function(x) !all(is.na(x)))]

  # If no variables are left after dropping, return NULL
  if (length(category_vars) == 0) {
    return(NULL)
  }

  # Create a subset of data with the dependent variable and independent variables
  regression_data <- data %>%
    select(Current_health_expenditure_per_capita_current_US, all_of(category_vars)) %>%
    na.omit()  # Remove rows with NA values

  # Check if there are enough rows to run the regression
  if (nrow(regression_data) == 0) {
    return(NULL)
  }

  # Create the regression formula
  formula <- as.formula(paste("Current_health_expenditure_per_capita_current_US ~", 
                              paste(category_vars, collapse = " + ")))

  # Run the regression
  regression_model <- lm(formula, data = regression_data)

  # Tidy up the regression results and add a "Significant" column
  tidy_results <- tidy(regression_model) %>%
    mutate(
      Category = category_name,
      Significant = ifelse(p.value <= 0.10, "YES", "NO")  # Add significance column
    )

  return(tidy_results)
}

# Function to run regression for all countries and a specific category
run_regression_all_countries <- function(category_vars, category_name) {
  results <- mydata %>%
    group_by(CountryName) %>%
    group_modify(~ {
      regression_results <- run_regression(.x, category_vars, category_name)
      if (!is.null(regression_results)) {
        return(regression_results)
      } else {
        return(tibble())  # Return empty tibble if no results
      }
    })

  return(results)
}

# Run regression for each category and create tables
economic_results <- run_regression_all_countries(variable_categories$Economic, "Economic")
health_results <- run_regression_all_countries(variable_categories$Health, "Health")
social_results <- run_regression_all_countries(variable_categories$Social, "Social")
government_results <- run_regression_all_countries(variable_categories$Government, "Government")
trade_results <- run_regression_all_countries(variable_categories$Trade, "Trade")

# Combine all results into one table
all_results <- bind_rows(
  economic_results,
  health_results,
  social_results,
  government_results,
  trade_results
)

# Ensure the "Significant" column has no NA values
all_results <- all_results %>%
  mutate(Significant = ifelse(is.na(Significant), "NO", Significant))  # Replace NA with "NO"

# Remove the CountryName column
all_results <- all_results %>%
  select(-CountryName)

# Rename the "term" column to "Predictor variables"
all_results <- all_results %>%
  rename(`Predictor variables` = term)

# Add category separators
all_results <- all_results %>%
  group_by(Category) %>%
  group_modify(~ {
    # Add a separator row before each category
    separator_row <- tibble(
      `Predictor variables` = paste("Independent variables from", .y$Category, "category"),
      estimate = NA, std.error = NA, statistic = NA, p.value = NA, Significant = NA
    )
    bind_rows(separator_row, .x)
  }) %>%
  ungroup()

# Create an advanced table using kableExtra
all_results %>%
  kable("html", caption = "Regression Results by Category") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%", height = "500px") %>%
  row_spec(which(all_results$Significant == "YES"), background = "#D9F2D9") %>%  # Highlight significant rows in green
  row_spec(which(all_results$`Predictor variables` %in% paste("Independent variables from", unique(all_results$Category), "category")), 
           bold = TRUE, background = "#FFA500") %>%  # Highlight separator rows in orange
  column_spec(1, bold = TRUE)  # Bold the "Predictor variables" column
```