### 2025-03-14 code/analysis-code/statistical-analysis.qmd
project part3

---
title: "statistical-analysis"
format: html
editor: visual
---

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      
library(kableExtra) 
library(webshot)
library(magick)
library(here)
library(modelsummary)
library(plm)
library(lmtest)
```

At first we will import the dataset

```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

We will now pick the most useful variables from the large pool of 1400+ variables in our dataset, and we will then categorize these variables based on the aspects they cover. For example, if there is avariable named "Annual GDP growth", we will categorize it as an "Economic" variable. Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable.

```{r}

# Define the dependent variable
dependent_var <- "Current_health_expenditure_per_capita_current_US"

# Select 50 variables that are likely to predict the dependent variable
selected_vars <- c(
  # Economic Factors
  "GDP_per_capita_current_US", "GNI_per_capita_Atlas_method_current_US", "GDP_growth_annual_",
  "Inflation_GDP_deflator_annual_", "Domestic_credit_provided_by_financial_sector__of_GDP",
  "Gross_capital_formation__of_GDP", "Exports_of_goods_and_services__of_GDP",
  "Imports_of_goods_and_services__of_GDP", "Foreign_direct_investment_net_inflows_BoP_current_US",
  "Industry_including_construction_value_added__of_GDP", "Agriculture_forestry_and_fishing_value_added__of_GDP",
  "Services_value_added__of_GDP", "Unemployment_total__of_total_labor_force_modeled_ILO_estimate",
  "Income_share_held_by_lowest_20", "Income_share_held_by_highest_20", "Poverty_headcount_ratio_at_215_a_day_2017_PPP__of_population",

  # Health Factors
  "Life_expectancy_at_birth_total_years", "Mortality_rate_under5_per_1000_live_births",
  "Prevalence_of_HIV_total__of_population_ages_1549", "Immunization_measles__of_children_ages_1223_months",
  "Births_attended_by_skilled_health_staff__of_total", "Contraceptive_prevalence_any_method__of_married_women_ages_1549",
  "Hospital_beds_per_1000_people", "Physicians_per_1000_people", "Prevalence_of_underweight_weight_for_age__of_children_under_5",
  "Adolescent_fertility_rate_births_per_1000_women_ages_1519", "Current_health_expenditure__of_GDP",
  "Domestic_general_government_health_expenditure__of_GDP",

  # Social Factors
  "Population_total", "Population_growth_annual_", "Population_density_people_per_sq_km_of_land_area",
  "Urban_population__of_total_population", "Rural_population__of_total_population",
  "Literacy_rate_adult_total__of_people_ages_15_and_above", "School_enrollment_primary__gross",
  "School_enrollment_secondary__gross", "School_enrollment_tertiary__gross", "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate",
  "Net_migration", "Refugee_population_by_country_or_territory_of_asylum",

  # Government Factors
  "Military_expenditure__of_GDP", "Tax_revenue__of_GDP", "Government_expenditure_on_education_total__of_GDP",
  "Government_Effectiveness_Estimate", "Rule_of_Law_Estimate", "Control_of_Corruption_Estimate",
  "Political_Stability_and_Absence_of_ViolenceTerrorism_Estimate", "Regulatory_Quality_Estimate",

  # Trade Factors
  "Trade__of_GDP", "Merchandise_trade__of_GDP", "Exports_of_goods_and_services_current_US",
  "Imports_of_goods_and_services_current_US", "Foreign_direct_investment_net_inflows_BoP_current_US",
  "External_debt_stocks_total_DOD_current_US", "Net_barter_terms_of_trade_index_2015__100"
)

# Add the dependent variable to the selected variables
selected_vars <- c(dependent_var, selected_vars)

# Subset the dataset with the selected variables
mydata_selected <- mydata %>%
  select(all_of(selected_vars))

# Categorize the selected variables
variable_categories <- list(
  Economic = c("GDP_per_capita_current_US", "GNI_per_capita_Atlas_method_current_US", "GDP_growth_annual_",
               "Inflation_GDP_deflator_annual_", "Domestic_credit_provided_by_financial_sector__of_GDP",
               "Gross_capital_formation__of_GDP", "Exports_of_goods_and_services__of_GDP",
               "Imports_of_goods_and_services__of_GDP", "Foreign_direct_investment_net_inflows_BoP_current_US",
               "Industry_including_construction_value_added__of_GDP", "Agriculture_forestry_and_fishing_value_added__of_GDP",
               "Services_value_added__of_GDP", "Unemployment_total__of_total_labor_force_modeled_ILO_estimate",
               "Income_share_held_by_lowest_20", "Income_share_held_by_highest_20", "Poverty_headcount_ratio_at_215_a_day_2017_PPP__of_population"),

  Health = c("Life_expectancy_at_birth_total_years", "Mortality_rate_under5_per_1000_live_births",
             "Prevalence_of_HIV_total__of_population_ages_1549", "Immunization_measles__of_children_ages_1223_months",
             "Births_attended_by_skilled_health_staff__of_total", "Contraceptive_prevalence_any_method__of_married_women_ages_1549",
             "Hospital_beds_per_1000_people", "Physicians_per_1000_people", "Prevalence_of_underweight_weight_for_age__of_children_under_5",
             "Adolescent_fertility_rate_births_per_1000_women_ages_1519", "Current_health_expenditure__of_GDP",
             "Domestic_general_government_health_expenditure__of_GDP"),

  Social = c("Population_total", "Population_growth_annual_", "Population_density_people_per_sq_km_of_land_area",
             "Urban_population__of_total_population", "Rural_population__of_total_population",
             "Literacy_rate_adult_total__of_people_ages_15_and_above", "School_enrollment_primary__gross",
             "School_enrollment_secondary__gross", "School_enrollment_tertiary__gross", "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate",
             "Net_migration", "Refugee_population_by_country_or_territory_of_asylum"),

  Government = c("Military_expenditure__of_GDP", "Tax_revenue__of_GDP", "Government_expenditure_on_education_total__of_GDP",
                 "Government_Effectiveness_Estimate", "Rule_of_Law_Estimate", "Control_of_Corruption_Estimate",
                 "Political_Stability_and_Absence_of_ViolenceTerrorism_Estimate", "Regulatory_Quality_Estimate"),

  Trade = c("Trade__of_GDP", "Merchandise_trade__of_GDP", "Exports_of_goods_and_services_current_US",
            "Imports_of_goods_and_services_current_US", "Foreign_direct_investment_net_inflows_BoP_current_US",
            "External_debt_stocks_total_DOD_current_US", "Net_barter_terms_of_trade_index_2015__100")
)

# Print the categorized variables
print(variable_categories)
```

Now that we have selected our variables, we will try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.

```{r}

# Convert all columns to numeric except CountryName, CountryCode, and SeriesCode
mydata <- mydata %>%
  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))

# Function to run regression for a specific category of independent variables
run_regression <- function(data, category_vars, category_name) {
  # Drop variables with all NA values within the category
  category_vars <- category_vars[sapply(data[category_vars], function(x) !all(is.na(x)))]

  # If no variables are left after dropping, return NULL
  if (length(category_vars) == 0) {
    return(NULL)
  }

  # Create a subset of data with the dependent variable and independent variables
  regression_data <- data %>%
    select(Current_health_expenditure_per_capita_current_US, all_of(category_vars)) %>%
    na.omit()  # Remove rows with NA values

  # Check if there are enough rows to run the regression
  if (nrow(regression_data) == 0) {
    return(NULL)
  }

  # Create the regression formula
  formula <- as.formula(paste("Current_health_expenditure_per_capita_current_US ~", 
                              paste(category_vars, collapse = " + ")))

  # Run the regression
  regression_model <- lm(formula, data = regression_data)

  # Tidy up the regression results and add a "Significant" column
  tidy_results <- tidy(regression_model) %>%
    mutate(
      Category = category_name,
      Significant = ifelse(p.value <= 0.10, "YES", "NO")  # Add significance column
    )

  return(tidy_results)
}

# Run regression for each category and create tables
economic_results <- run_regression(mydata, variable_categories$Economic, "Economic")
health_results <- run_regression(mydata, variable_categories$Health, "Health")
social_results <- run_regression(mydata, variable_categories$Social, "Social")
government_results <- run_regression(mydata, variable_categories$Government, "Government")
trade_results <- run_regression(mydata, variable_categories$Trade, "Trade")

# Combine all results into one table
all_results <- bind_rows(
  economic_results,
  health_results,
  social_results,
  government_results,
  trade_results
)

# Ensure the "Significant" column has no NA values
all_results <- all_results %>%
  mutate(Significant = ifelse(is.na(Significant), "NO", Significant))  # Replace NA with "NO"

# Rename the "term" column to "Predictor variables"
all_results <- all_results %>%
  rename(`Predictor variables` = term)

# Add category separators
all_results <- all_results %>%
  group_by(Category) %>%
  group_modify(~ {
    # Add a separator row before each category
    separator_row <- tibble(
      `Predictor variables` = paste("Independent variables from", .y$Category, "category"),
      estimate = NA, std.error = NA, statistic = NA, p.value = NA, Significant = NA
    )
    bind_rows(separator_row, .x)
  }) %>%
  ungroup()

# Create an advanced table using kableExtra
table_html <- all_results %>%
  kable("html", caption = "Regression Results by Category") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%", height = "500px") %>%
  row_spec(which(all_results$Significant == "YES"), background = "#D9F2D9") %>%  # Highlight significant rows in green
  row_spec(which(all_results$`Predictor variables` %in% paste("Independent variables from", unique(all_results$Category), "category")), 
           bold = TRUE, background = "#FFA500") %>%  # Highlight separator rows in orange
  column_spec(1, bold = TRUE)  # Bold the "Predictor variables" column

# Display the table
table_html
```

In the above table, we can observe that a lot of factors affect the dependent variable and a lot of them don't. We can also observe that a lot of our variables have missing values make it unavailable for the regression estimation.

Now, we will save our regression result is a HTML table so that we can use to present it in our manuscript.

```{r}
# Save the table as a PDF file
html_file <- here("results", "tables", "regression_results.html")
save_kable(table_html, file = html_file)

```

Now that we have tested some variables in our model across different categories, we will pick the most useful variable and add them into our preliminary model. We will also add other variable that were not present in the above table. Variable selection is a continuous process and the variables selected in the preliminary model may also be dropped later to make space for new variables that may appear as important.

```{r}

model <- lm(Current_health_expenditure_per_capita_current_US ~ 
              GDP_growth_annual_ + 
              Income_share_held_by_lowest_20 + 
              Control_of_Corruption_Estimate + 
              Life_expectancy_at_birth_total_years + 
Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
              Literacy_rate_adult_total__of_people_ages_15_and_above + 
              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+ 
  Access_to_electricity_rural__of_rural_population +
  Age_dependency_ratio__of_workingage_population
  , 
            data = mydata)

# Save the regression results to an HTML file
modelsummary(
  model,
  output = here("results", "tables", "preliminary_regression_results.html"),
  title = "First Attempt at Regression Model",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  ),
  encoding = "UTF-8"  # Ensure UTF-8 encoding
)
```

The variables have a significant impact on the dependent variable as expected. But, as this is a time series dataset, we must also test for autocorrelation. There are many other tests as well which can be performed on time-series models such ADF test for unit root, and complex models such as ARCH and GARCH can also be used. However, we will only perform simple procedures and try to build a simple model.

# Autocorrelation test

Below, we will perform the autocorrelation test.

```{r}
# Perform Durbin-Watson test
dw_test <- dwtest(model)
print(dw_test)
```

Unfortunately, the Durbin-Watson test is significant, and our model shows presence of autocorrelation.

We will try to use a lagged model to solve the issue of autocorrelation.

```{r}
# Create a lagged version of the dependent variable
mydata$Current_health_expenditure_per_capita_current_US_lag1 <- lag(mydata$Current_health_expenditure_per_capita_current_US, 1)

mydata$GDP_growth_annual_lag1 <- lag(mydata$GDP_growth_annual_, 1)

# Refit the model with lagged variables
model_lagged <- lm(Current_health_expenditure_per_capita_current_US ~ 
                     GDP_growth_annual_ + GDP_growth_annual_lag1 +
                     Income_share_held_by_lowest_20 + 
                     Control_of_Corruption_Estimate + 
                     Life_expectancy_at_birth_total_years + 
                     Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
                     Literacy_rate_adult_total__of_people_ages_15_and_above + 
                     Trade__of_GDP + 
                     Foreign_direct_investment_net_inflows_BoP_current_US + 
                     Access_to_electricity_rural__of_rural_population +
                     Age_dependency_ratio__of_workingage_population +
                     Current_health_expenditure_per_capita_current_US_lag1,  # Include lagged dependent variable
                   data = mydata)

# autocorrelation test

dw_test_lagged <- dwtest(model_lagged)
print(dw_test_lagged)
```

Unfortunately, the DW test is still significant, and lagged model also shows auto correlation.

We will try to use a country fixed effect to account for the all time-invariant differences between countries (e.g., cultural, institutional, or geographical factors). Although these country fixed effect is not a solution to the autocorrelation problem in itself, we believe that the new model specification will show a different property than our previous two models with ordinary pooled OLS and lagged model.

```{r}
# Fit a fixed effects model using lm() with dummy variables
fixed_model_lm <- lm(Current_health_expenditure_per_capita_current_US ~ 
                       GDP_growth_annual_ + 
                       Income_share_held_by_lowest_20 + 
                       Control_of_Corruption_Estimate + 
                       Life_expectancy_at_birth_total_years + 
   Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate +
                       Literacy_rate_adult_total__of_people_ages_15_and_above + 
                       Trade__of_GDP + 
                       Foreign_direct_investment_net_inflows_BoP_current_US + 
                       Access_to_electricity_rural__of_rural_population +
                       Age_dependency_ratio__of_workingage_population + 
                       factor(CountryName),  # Include dummy variables for each entity
                     data = mydata)

#autocorrelation test

dw_test_lagged <- dwtest(fixed_model_lm)
print(dw_test_lagged)
```

Thankfully, our country fixed effect model no longer shows autocorrelation. We will now be able to build upon this model for our further analysis.