### 2025-04-03 code/robustness-code/robustness-code.qmd

---
title: "robustness-check"
format: html
editor: visual
---

```{r}
# Load required packages
library(purrr)    # For map_df()
library(dplyr)    # For data manipulation
library(timetk)   # For time_series_cv()
library(rsample)  # For analysis() and assessment()
library(here)
library(modelsummary)
library(tidyr)
library(ranger)

```

Importing train/test dataset.

```{r}
train <- readRDS(here("data", "processed-data", "train-data.rds"))
test <- readRDS(here("data", "processed-data", "test-data.rds"))
```

testing the results

```{r}
# preliminary regression model

model <- lm(Current_health_expenditure_per_capita_current_US ~ 
              GDP_growth_annual_ + 
              Income_share_held_by_lowest_20 + 
              Control_of_Corruption_Estimate + 
              Life_expectancy_at_birth_total_years + 
Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+ 
  Access_to_electricity_rural__of_rural_population +
  Age_dependency_ratio__of_workingage_population
  , 
            data = test)

modelsummary(
  model,
  output = here("results", "tables", "testresult.png"),
  title = "This table represents regresssion results from the test data",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

```

# cross validation

```{r}

# 1. Define your model formula (as you originally specified)
model_formula <- Current_health_expenditure_per_capita_current_US ~ 
  GDP_growth_annual_ + 
  Income_share_held_by_lowest_20 + 
  Control_of_Corruption_Estimate + 
  Life_expectancy_at_birth_total_years + 
  Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
  Trade__of_GDP + 
  Foreign_direct_investment_net_inflows_BoP_current_US + 
  Access_to_electricity_rural__of_rural_population +
  Age_dependency_ratio__of_workingage_population

# 2. Create time-series CV splits
cv_splits <- time_series_cv(
  data = train,
  date_var = Year,
  initial = 5,  # 5-year initial window
  assess = 1,   # 1-year test window
  skip = 1,     # Move forward 1 year each split
  slice_limit = 3
)
```

```{r}

set.seed(123)

cv <- function(splits, model_formula) {
  map_df(splits, function(split) {
    train_cv <- analysis(split)
    test_cv <- assessment(split)

    # Train model
    model <- lm(model_formula, data = train_cv)

    # Get predictions and actuals
    preds <- predict(model, newdata = test_cv)
    actuals <- test_cv$Current_health_expenditure_per_capita_current_US

    # Convert to numeric safely
    preds_num <- suppressWarnings(as.numeric(preds))
    actuals_num <- suppressWarnings(as.numeric(actuals))

    # Calculate RMSE with error handling
    if(length(preds_num) == 0 || length(actuals_num) == 0) {
      rmse_val <- NA_real_
      status <- "Empty predictions/actuals"
    } else if(any(is.na(actuals_num))) {
      valid <- !is.na(actuals_num)
      rmse_val <- sqrt(mean((preds_num[valid] - actuals_num[valid])^2, na.rm = TRUE))
      status <- "NA in actuals"
    } else {
      rmse_val <- sqrt(mean((preds_num - actuals_num)^2, na.rm = TRUE))
      status <- "Success"
    }

    data.frame(
      Train_Years = paste(min(train_cv$Year), max(train_cv$Year), sep = "-"),
      Test_Year = max(test_cv$Year),
      RMSE = rmse_val,
      stringsAsFactors = FALSE
    )
  })
}

# Run the analysis
results <- cv(cv_splits$splits, model_formula)
print(results)
```

```{r}
# View summary of CV results
cv_summary <- results %>%
  summarise(
    Mean_RMSE = mean(RMSE, na.rm = TRUE),
    SD_RMSE = sd(RMSE, na.rm = TRUE),
  )

print(cv_summary)
```

# Random Forest

```{r}
rf_cv <- function(splits, model_formula) {
  map_df(splits, function(split) {
    train_cv <- analysis(split)
    test_cv <- assessment(split)

    # 1. Get target variable from formula
    target_var <- all.vars(model_formula)[1]

    # 2. Check if Year column exists (case-insensitive)
    year_col <- grep("^year$", names(train_cv), ignore.case = TRUE, value = TRUE)

    # 3. Select required variables (including Year if it exists)
    required_vars <- unique(c(all.vars(model_formula), year_col))
    train_cv <- train_cv %>% 
      select(any_of(required_vars)) %>% 
      drop_na(all_of(target_var))

    test_cv <- test_cv %>% 
      select(any_of(required_vars)) %>% 
      drop_na(all_of(target_var))

    # 4. Get year information
    get_years <- function(df) {
      if (length(year_col) > 0) {
        list(
          min_year = min(df[[year_col]], na.rm = TRUE),
          max_year = max(df[[year_col]], na.rm = TRUE)
        )
      } else {
        list(min_year = NA_real_, max_year = NA_real_)
      }
    }

    train_years <- get_years(train_cv)
    test_year <- get_years(test_cv)$max_year

    # 5. Calculate mtry
    n_predictors <- length(setdiff(required_vars, c(target_var, year_col)))
    mtry_value <- max(1, min(floor(n_predictors / 3), n_predictors))

    # 6. Train model
    set.seed(123)
    rf_model <- ranger(
      formula = model_formula,
      data = train_cv,
      num.trees = 500,
      mtry = mtry_value,
      importance = 'permutation',
      seed = 123
    )

    # 7. Calculate RMSE
    preds <- predict(rf_model, data = test_cv)$predictions
    actuals <- pull(test_cv, target_var)
    rmse_val <- sqrt(mean((preds - actuals)^2, na.rm = TRUE))

    # 8. Return results
    tibble(
      Train_Years = if (!is.na(train_years$min_year)) {
        paste(train_years$min_year, train_years$max_year, sep = "-")
      } else {NA_character_},
      Test_Year = test_year,
      RMSE = rmse_val,
      Model = "Random Forest",
      Predictors_Used = n_predictors,
      mtry_Used = mtry_value,
      Year_Column_Found = ifelse(length(year_col) > 0, year_col, "Not found")
    )
  })
}
```

```{r}
rf_results <- rf_cv(cv_splits$splits, model_formula)
print(rf_results)
```

# Random Forest on test dataset

```{r}
library(tidyverse)
library(ranger)

# Prepare training data (remove NA in target variable)
target_var <- all.vars(model_formula)[1]
train_processed <- train %>%
  drop_na(all_of(target_var))

# Calculate appropriate mtry
n_predictors <- length(all.vars(model_formula)) - 1
mtry_value <- max(1, min(floor(n_predictors / 3), n_predictors))

# Train Random Forest model
set.seed(123)
rf_model <- ranger(
  formula = model_formula,
  data = train_processed,
  num.trees = 500,
  mtry = mtry_value,
  importance = 'permutation',
  seed = 123
)
```

```{r}
# Ensure test data has same structure as training data
test_processed <- test %>%
  select(any_of(names(train_processed))) %>%  # Keep only columns present in training
  drop_na(all_of(target_var))  # Remove rows with NA in target
```

```{r}
# Generate predictions
test_predictions <- predict(rf_model, data = test_processed)$predictions

# Create results dataframe
test_results <- test_processed %>%
  mutate(Predicted = test_predictions,
         Residual = .data[[target_var]] - Predicted)
```

```{r}
# Calculate metrics
performance_metrics <- test_results %>%
  summarise(
    RMSE = sqrt(mean(Residual^2, na.rm = TRUE)),
    MAE = mean(abs(Residual), na.rm = TRUE),
    R2 = 1 - (sum(Residual^2) / sum((.data[[target_var]] - mean(.data[[target_var]]))^2))
  )

# Print metrics
cat("\nTest Set Performance:\n")
print(performance_metrics)

# View predictions vs actual
head(test_results %>% select(all_of(target_var), Predicted, Residual))
```