### 2025-04-24 review-zuber/project-review-template.qmd
Part 5 review_ Mohammed Zuber

---
title: Project Review Template 
author: Mohammed Zuber
date: 4/23/2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Factors affecting health sector expenditure in OECD countries

Name of project author(s): Prasanga Paudel

Name of project reviewer: Mohammed Zuber

# Score 
Overall Score: 3 (Sufficient). 
The project is complete, detailed, and well-documented. It includes a comprehensive analysis using both statistical and machine learning approaches, with thorough explanation and reproducibility. Minor improvements are suggested, but they do not detract from the overall high quality of the work.
# Instructions
## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background provides a clear rationale for studying health expenditure with references to prior studies. The objective to understand determinants such as GDP growth, corruption, and demographic variables is clearly articulated. However, a more detailed summary of the gaps in previous research and how this study addresses them would strengthen the motivation.

### Summary assessment 
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
The research questions are explicitly stated and directly linked to the outcome variable (health expenditure per capita). The two-part objective — assessing key factors and their magnitude of influence — is well aligned with the analytical plan.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
The data source (World Bank via OECD) is clearly identified, and the use of 462 observations from 38 countries is explained well. The GitHub repository is mentioned as the location for preprocessing steps, which is helpful. However, a codebook or variable dictionary is missing from the project folder, which makes it harder to understand the meaning and units of the 60+ variables used. Including even a brief data dictionary or a summary table with variable names and definitions would enhance transparency and usability.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments
Exploratory analysis is well-executed with appropriate figures (e.g., trends, correlation plots) and justified variable selection. The well documented comments on how much time the code takes to run is commendable.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The OLS regression is appropriate and well-justified for panel data. The inclusion of cross-validation and RF as robustness checks adds value.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
Figures and tables are informative and generally clear. However, a few visuals could benefit from better axis labeling and reduced clutter. The narrative does a good job walking the reader through the findings.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion effectively synthesizes the key findings and connects them to existing literature, reinforcing the credibility of the results. The implications for public policy and international resource allocation are thoughtfully addressed. However, the discussion lacks a clear summary of the study’s strengths and limitations. For instance, a major strength is the comprehensive modeling approach, including OLS and Random Forest with validation across datasets. A notable limitation is the potential overfitting in the RF model and the absence of detail on how missing data and variable collinearity were handled. Including these would provide a more balanced and critical reflection on the results.

### Summary assessment
* minor parts wrong, missing or unclear

## Further comments
none

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project folder is clearly organized, with appropriately named files and a logical directory structure. There are no unnecessary or junk files, and the layout makes it easy to follow the workflow from raw data to final analysis and figures. The repository reflects good practices in reproducible research and is easy to navigate.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project is well documented, with clear and consistent comments throughout the code. Each line of code is annotated to explain its purpose, making the analysis easy to follow. The Rmd files guide the reader through the workflow step by step, and the folder structure further supports understanding. Overall, the documentation reflects thoughtful effort to ensure transparency and reproducibility.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The project is fully reproducible, with all code executing smoothly and generating the expected outputs, including figures and tables. The analysis runs without any need for manual intervention, and all necessary data files and scripts are included in the repository. This reflects strong adherence to reproducible research standards.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The project demonstrates a strong level of thoroughness. Multiple modeling approaches were considered — including OLS and Random Forest — and each was tested using cross-validation and out-of-sample performance.

### Summary assessment
* strong level of thorougness

## Further comments

Overall, Great Job!