### 2025-05-03 products/manuscript/Manuscript.qmd
#final edits

---
title: "Factors affecting health sector expenditure in OECD countries"
author: "Prasanga Paudel"
format:
  html:
    toc: false
    embed-resources: true
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/vancouver-author-date.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(here)
library(readxl)
library(writexl)
library(magrittr)
library(patchwork)
library(kableExtra)
```

**Authors**

-   Prasanga Paudel$^{1}$ (ORCID: 0000-0000-1234-5678)

**Author affiliations**

1.  College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.

$\dagger$ Disclaimer: The opinions expressed in this article are the author's own and don't reflect their employer.

{{< pagebreak >}}

<strong>Abstract</strong>

_Understanding the determinants of health expenditures is a critical job in order to optimize the healthcare systems and policymaking. This study utilizes the Ordinary Least Squares method to analyze the factors affecting per capita health expenditure of a country among the OECD nations during 2010 to 2020. Our study identifies that GDP growth, control of corruption, and  unemployment rate among the youth in the labor force have a significant and an important role in determining the amount of money spent on health sector by the country, while Foreign Direct Investment exhibits a significant but low magnitude impact. Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration showed no statistical significant impact on the health expenditure. We also applied Random Forest model to explain the factors affecting per capita health expenditure of a country. The results of both the OLS model and Random forest was tested using cross-validation and train/test technique, and we found that our OLS model shows a consistent result across all tests. Although Random Model was found to have a better explanation capability, the inconsistensies raised a question on it's validity. Our findings highlight the importance of economic and institutional factors in shaping the health expenditure and offer insights for policimakers aiming to enhance healthcare accessibility and efficiency. _

{{< pagebreak >}}

# Introduction

## Health expenditure as an indicator

Health expenditure is an important indicator of a country's committment to citizen's well-being and efficiency of it's healthcare systems. Among the OECD (Organisation for Economic Co-operation and Development) countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 USD [@2021].
Understanding the factors that drive these differences is important for policymakers to optimize healthcare delivery and improving healthcare access. While GDP had been one of the primary determinant of health expending [@Newhouse1977], recent studies suggest that institutional, demographic, and socio-economic factors may also play a crucial role in determination of health expenditure among different countries [@Ziegler2017]. The relationship between health expenditure and their determinants because particularly relevant in the context of recent global challenges. The recent COVID-19 pandemic also exposed the vulnerabilities in global healthcare systems with countries showing high pre-pandemic health expenditure generally demonstrating greater resilience [@WHO2022]

We will try to analyze the most influential factors affecting the expenditure in health sector by different countries. The predictor variables will include factors ranging from education, mortality, population distribution among rural and urban spaces, and so on. We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors for this panel data.

We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure. Some research have such as [@Nkemgha2020] studied the link among public health expenditure, private health expenditure and life expectancy in Cameroon. [@CHRISTOPOULOS2020195] using panel data analyzed the fiscal effects of health expenditures on health outcomes and found a significant impact of healthcare expenditures on increasing revenue. In our analysis we will try to follow OLS multiple regression model. We will also use methodological inventions in comparing different models primarily OLS regression and Random Forest to address the recent suggestions in analytical apporaches in health economics [@ParkerLue2020]

## Data source

This is a time series data based on 38 member countries from the Organisation for Economic Co-operation and Development (OECD). This organization is a unique forum with governments from 37 democracies collaborate to develop policy standards to promote sustainable economic growth. However, the data obtained from this dataset is crude and needed to be treated before attempting analysis based on it. The details about the data cleaning process is available at [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project).

The original data (2010-2020) in it's raw form was obtained from World Bank. However, the dataset needed processing, and some modification. The final dataset that is eligible for our use contains 462 observations and 1496 variables for the OECD countries. The variables include information on various aspect of a country's development indicators, such as information on education, health, population distribution, trade, energy access, GDP, infrastructure, market, corruption, investments, and so on. However, we will be using 80% of the dataset for our whole analysis and we will test the robustness of the results on the un-used 20% at the end of the analysis.

## Objective of the research

The expenditure in health sector is one of the important factor that helps in determining the efficiency of health sector. We can relate the expenditure with different national development indicators and hope to find important connections between them. This will help government and international institutions to make specific interventions through their policy to enhance their health sector. We can also find out how countries are currently operating and which factor has helped or negatively affected their health sector.

The main objective of this research is to:

1. To analyze the factors that affect the expenditure in health sector among OECD countries.

2. To analyze the magnitude of the influence of these factors on health expenditure.

{{< pagebreak >}}

# Methods

We will be following a panel data time series regression model. A general form of the OLS model for time series can be represented as:

   $y_{it}$ = $b_0$ + $b_1$$X_{it1}$ + $b_2$$X_{it2}$ + $b_3$$X_{it3}$ + $b_4$$X_{it4}$ + $b_5$$X_{it5}$ + $e_{it}$

where $e_{it}$ is the error term. $X_{itk}$ is the $k^{th}$ independent variable ( k= 1 to 5 in the above equation but our research will likely have k>5), $y_{it}$ is the dependent variable, $b_0$ is the intercept, and $b_k$ is the estimated parameter for the $k^{th}$ independent variable.

{{< pagebreak >}}

# Results

In this section, we will we will explore the data and perform analysis based on ordinary least squares and random forest model.

## Descriptive analysis

Here, we will now use a combination of tables and figues to have a look into the data.

As this is a time series data, it is quite difficult to visualize all the variables across the years for all the countries. So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years or the average trend.

### Trends of Development Indicators

In @fig-development-indicators-trends, we will explore differenct variables of interest by observing the average trends across the countries. We can observe that average life expectancy has been increasing consistently through 2010 to 2020. The average physicians per 1000 people shows an upward trend. However, hospital beds show a decreasing trend. The average government effectiveness seems to be dereasing across the countries. The unemployment rate in the OECD countries seems to be consistently decreasing. Finally, if we observe the average Gross Domestic Product(GDP) , we can see that the GDP of OECD countries are somewhat consistent except for the last year 2018.

```{r}
#| label: fig-development-indicators-trends
#| fig-cap: "Trends of Development Indicators (average across countries)"
#| echo: FALSE
knitr::include_graphics(here("results","figures","development_indicators_trends.png"))
```

### Trends of health expenditure

@fig-combined-health-expenditure-plot shows how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" and "Current Per Capita Health Expenditure (2010-2020)" are distributed over the years for top 5 OECD countries. We can observe that among the two factors, Health Expenditure as a percentage of GDP is relatively more consistent across the years. Per capita health expenditure across the countries, on the other hand, shows variation across times. It is evident that per capaita health expenditure is more sensitive to external factors and shows immediate response through each years. Our objective is to explain this variation in expenditure.

```{r}
#| label: fig-combined-health-expenditure-plot
#| fig-cap: "Health expenditure across top 5 countries"
#| echo: FALSE
knitr::include_graphics(here("results","figures","combined_health_expenditure_plots.png"))
```

## Basic statistical analysis

### Preparing a model

To help finalize our model, we first tried to look at the relation between our dependent variable "Health Expenditure Per Capita $USD" with other variables from a specific category. Specifically, we picked the most useful variables from the large pool of 1400+ variables in our dataset, and we then categorized these variables based on the aspects they cover. For example, if there is a variable named "Annual GDP growth", we will categorize it as an "Economic" variable. Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable

We then ran our initial regression with "Health Expenditure Per Capita (USD)" as dependent variable, and variables from each of the category as independent variables in separate regressions. The motivation here was to find the most influential factor from each category. The regression results for this preliminary test can be found in  [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project). After observing the relation of almost 60 important variables with our variable of interest "Health Expenditure Per Capita $USD", we filtered few variables worth pursuing, and we will discuss them in the sections that will come.

### Observing the relationship with dependent variable

@fig-correlation-p1 shows the correlation of per capita health expenditure with different variables for the year 2010. This fugure shows the pattern present in the dataset for the year 2010, with all the countries treated as cross-sectional units. We are only using the information from one year, year 2010, because visualization across years can be quite complex as we would also have to consider for the trends of each country across years.

```{r}
#| label: fig-correlation-p1
#| fig-cap: "Correlation based on 2010 data"
#| echo: FALSE
#| out.width: "100%"
knitr::include_graphics(here("results","figures","correlation_plot1.png"))
```

@fig-correlation-p1 shows that income share held by lowest 20% of the population is positively correlated with the per capita health expenditure. Similary, estimate of control of corruption and life expectancy at birth also show positive correlation. Annual GDP growth on the other hand shows a low but negative correlation.

We will check few other relation in @fig-correlation-p2.

```{r}
#| label: fig-correlation-p2
#| fig-cap: "Additional correlation based on 2010 data"
#| echo: FALSE
#| out.width: "100%"
knitr::include_graphics(here("results","figures","correlation_plot2.png"))
```

It can be seen that most of the variables seem to have a moderate and positive correlation with our variable of interest. Although some variables lack observations, we must recall that this is just for the year 2010, and the whole dataset might have a different distribution pattern.

### Proposed Model for the research {#model}

Based on the informations obtained from our preliminary results, we have selected variable that would best explain our dependent variable. The model is as follow:

Current Health Expenditure Per Capita (\$USD) = $b_0$ + $b_1$ Annual GDP growth + $b_2$ Income share held by lowest 20% + $b_3$ Control of corruption + $b_4$ Life expectancy at birth + $b_5$ Unemployed youth population as a percentage of total labor force + $b_6$ Trade as a percentage of GDP + $b_7$ FDI net inflow + $b_8$ Access to electricity (rural) + $b_{9}$ Age dependency ratio

## Full analysis

### Regression Results

We ran the regression based on our primary model described in the section [ 5.2.3](#model) of this article. The results obtained from the regression analysis are presented by @tbl-primary-regression-results.

```{r}
#| label: tbl-primary-regression-results
#| tbl-cap: "Primary Regression Results"
#| echo: FALSE
#| out.width: "70%"
knitr::include_graphics(here("results", "tables", "primary_regression_results.png"))
```

The result shows that annual GDP growth of a country has statistically significant negative impact on the per capita health expenditure of the country at 1% level of significance. Specifically, one percentage oint growth in annual GDP decreases the per capita health expenditure by 117.468 US dollars. Similarly, the estimate for the control of corruption in a country has a statistically significant negative effect at 1% level of significance. Specifically, one unit increase in control increases the per capita health expenditure by 1822 US dollars. The level of unemployment among youth population out of the total labor force population had a statistically significant negative effect at 1% level of significance. To be precise, one percent increase in unemployment rate among youth decreases the per capita health expenditure by 25 US dollars. Similarly, foreign direct investment (FDI) had a positive but very low magnitude effect on the health expenditure at 1% level of significance. Life expectancy at birth had a statistically significant positive effect on the health expenditure at 1% significance level, with increase in health expectancy by one year resulting in an increase in per capita health expenditure by 208 US dollars. Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.

The model showed a "good" descriptive performance based on it's adjusted R-squared value of 74.3%. The RMSE value for our model (RMSE=1213) is also quite lower compared to the NULL Model which showed a RMSE value of 2364.

#### Autocorrelation Test

We performed the Durbin Watson autocorrelation test for our model. The results from the DW test, presented by @tbl-DW-test-results, confirms the absence of autocorrelation, therefore we can move forward with our model.

```{r}
#| label: tbl-DW-test-results
#| tbl-cap: "Durbin Watson Test Results"
#| echo: FALSE
#| out.width: "30%"
knitr::include_graphics(here("results", "figures", "durbin_watson_test_results.png"))
```

#### ADF unit root test

We also performed the Fisher-type Augmented Dickey Fuller (ADF) test for presence of unit root. The results confirmed that the not all series in our panel data are non-stationary.

```{r}
#| label: tbl-unit-root-results
#| tbl-cap: "ADF Unit Root Test Results"
#| echo: FALSE

htmltools::tags$iframe(
  src = here("results", "tables", "unit_root_results.html"),
  width = "100%",
  height = "150px",
  scrolling = "auto",
  frameborder = "0"
)
```

Since there is no significant presence of correlation and unit root, we will claim our model to be a valid model. Next, we will test the robustness of our OLS results throuh various machine learning techniques.

#### Cross-Validation of OLS results

In this section, we will observe how our model will perform across different subsets of datasets within our train dataset.As this is time-series dataset, we will consider the fact that the values across different years are correlated with each other. We, therefore, use three different subset of years with five years within each subset as training data within the train dataset. We will use the year after the training dataset as a testing dataset. Therefore, we ended up with three subsets of training data and three testing data.

```{r}
#| label: tbl-OLS-CV-results-full
#| tbl-cap: "OLS cross-validation results "
#| echo: FALSE

htmltools::tags$iframe(
  src = here("results", "tables", "OLS-CV-results-full.html"),
  width = "100%",
  height = "200px",
  scrolling = "auto",
  frameborder = "0"
)
```

It can be observed that the RMSE values across the different combinations of train/test dataset show a somewhat consistent result. The RMSE value for the primary dataset was 1213. The results from CV, as observed in @tbl-ols-cv-results, show even less RMSE in some cases. This is a good thing as our model performs consistently better in cross-validation.

```{r}
#| label: tbl-ols-cv-results
#| tbl-cap: "OLS cross-validation results summary"
#| echo: FALSE

htmltools::tags$div(
  style = "display: flex; justify-content: center;",
  htmltools::tags$iframe(
    src = here("results", "tables", "OLS-CV-results.html"),
    width = "40%",
    height = "110px",
    scrolling = "auto",
    frameborder = "20"
  )
)
```

If we observe the mean RMSE across the three cross-validation subsets, the mean is lower than our primary dataset with a small standard deviation.

### Random Forest model as an Alternative

Next, we will perform our analyses with a different technique other than ordinary least squares method. We will use a Random Forest (RF) model to explain per capita health expenditure of a country. we will use a tree size of 300, and mtry value equal to 3. The motivation behind using a random forest model is to check if we are being able to capture the relation between per capita health expenditure and our predictors variable well enough using the OLS model. The non-linear effects, if present, would be captured better by the RF model.

#### Random Forest Results

@tbl-rf-importance-results represents the importance factor obtained from the random forest model. It can be observed that the variables that were found to have a significant and high impact on the per capita health expenditure in the OLS model also show a high importance in the random forest model as well, although results are different from OLS as well.

```{r}
#| label: tbl-rf-importance-results
#| tbl-cap: "RF importance plot"
#| echo: FALSE
#| out.width: "100%"
knitr::include_graphics(here("results", "figures", "rf_importance_plot.png"))
```

#### Performance of RF model

The performance of our random forest model is quite exciting. The R-squared value obtained is quite high (0.988. The RMSE value is also 288, which is almost 1000 less than the RMSE value (1213) obtained from our OLS model. This shows that RF can explain our data way more appropriately than an OLS model.

However, we need to test if this result is consistent across various dataset.

```{r}
#| label: tbl-rf-train-perform
#| tbl-cap: "Random Forest TRAIN dataset performance"
#| echo: FALSE

htmltools::tags$iframe(
  src = here("results", "tables", "rf_train_performance.html"),
  width = "100%",
  height = "200px",
  scrolling = "auto",
  frameborder = "0"
)
```

#### Cross-validation of RF model

In this section, we will try to see how our RF performs across various cross-validation datasets. We will perform cv based on the same combination of datasets we used for our OLS model. There are three different subsets of training dataset with five years in each dataset. Also, there will be a one year test dataset for the year immediately coming after the last year of testing dataset.

```{r}
#| label: tbl-rf-cv-results
#| tbl-cap: "Random Forest CV Results"
#| echo: FALSE

htmltools::tags$iframe(
  src = here("results", "tables", "rf_cv_results.html"),
  width = "100%",
  height = "300px",
  scrolling = "auto",
  frameborder = "0"
)
```

The results obtained from the cross-validation as observed in @tbl-rf-cv-results show that the RMSE values have more than doubled in the cross-validation sets. Infact, the highest RMSE value observed in cv is as high as 2.6 times more than the RMSE obtained in the primary train dataset.

```{r}
#| label: tbl-rf-cv-summary
#| tbl-cap: "Random Forest CV Summary"
#| echo: FALSE

htmltools::div(
  style = "text-align: center;",
  htmltools::tags$iframe(
    src = here("results", "tables", "rf_cv_summary.html"),
    width = "100%",
    height = "120px",
    scrolling = "auto",
    frameborder = "0"
  )
)
```

The mean RMSE, 644, despite showing a standard deviation comparable to the OLS model, has a significantly higher value compared to RMSE (288) obtained with the train dataset. This means that our RF model is performing poorly in the case of different subsections of datasets.

### Visual inspection of the two models in train dataset

Finally, in this section, we will try to visually observe the performance of the two models (OLS and Random Forest) in our train dataset that we discussed and worked on till now.

```{r}
#| label: fig-pred-vs-obs-train
#| fig-cap: "Predicted vs Observed values based on TRAIN dataset"
#| echo: FALSE
knitr::include_graphics(here("results","figures","pred_vs_obs_train.png"))
```

It can be observed in @fig-pred-vs-obs-train that the Random Forest clearly out-performs the OLS model in correctly predicting the outcomes. The datapoints for OLS regression are comparitively very hapazardly distributed compared to the datapoints of RF model.

However, we should remember that although the RF model had an outstanding performance in our primary train dataset with RMSE almost 1000 less than RMSE observed in the OLS model, the cross-validation results showed significant fluctuations, due to which robustness of our RF model can be questioned.

### Test dataset

#### Perfrormance of OLS on test dataset

We also performed the analysis on test dataset. The regression result obtained for the testing dataset is as follows.

```{r}
#| label: tbl-test-regression-results
#| tbl-cap: "Regression Results for the test dataset"
#| echo: FALSE
#| out.width: "70%"
knitr::include_graphics(here("results", "tables", "testresult.png"))
```

We can observe that the variables in general show some common pattern between the two datasets. The difference is acceptable as they are based on entirely different time frames. If we observe the model performance, it is evident that the model performance is weaker than obtined in the train dataset. The RMSE has increased by 383 from 1213 to 1596, however this is still 768 lower than the NULL model. The adjusted R-squared in our test dataset has also dropped by 12.3% to 62%.

#### Performance of Random Forest on test dataset

We also ran our Random Forest model with the train test dataset. The performance metrics obtained from the model are presented in @tbl-rf-test-perform.

```{r}
#| label: tbl-rf-test-perform
#| tbl-cap: "Random Forest TEST dataset performance"
#| echo: FALSE

htmltools::tags$iframe(
  src = here("results", "tables", "rf_test_performance.html"),
  width = "80%",
  height = "250px",
  scrolling = "auto",
  frameborder = "0"
)
```

We can observe that the RMSE value has again largely increased. The RMSE from test dataset is now as large as 4.6 times more than the RMSE obtained in the primary train dataset. This value is also double than the mean RMSE obtained from cross-validations. Therefore, there is a high suspicion that our our primary RF model sufferes from overfitting, although it performs relatively better than our OLS model.

But, as of now, as our OLS model showed a consistent and robust result across all our datasets and robustness tests, we can conclude that OLS model should be preferred over the Random Forest model despite RF performs better than OLS. Our conclusion is that the RF model can not be trusted because of it's inconsistency and suspicious results.

#### Visual inspection in test dataset

Finally, we will visually observe the performance of both the models in the test dataset. @fig-pred-vs-obs-test shows the distribution of observed values and predicted values using test dataset for both the models.

```{r}
#| label: fig-pred-vs-obs-test
#| fig-cap: "Predicted vs Observed values based on TEST dataset"
#| echo: FALSE
knitr::include_graphics(here("results","figures","pred_vs_obs_test.png"))
```

As expected based on the RMSE values, the RF model does perform better than our OLS model, However, due to inconsistency, we will prefer our OLS model over the RF model.

One thing that should be noted is that the RF model should be optimized and formulated in an efficient manner because it clearly outperformed the OLS model in every comparision across different datasets despite the inconsistency.

{{< pagebreak >}}

# Discussion

Our results show that annual GDP growth, control of corruption, life expectancy at birth, unemployment rate, and FDI all have a statistically significant effect on the per capita health expenditure of a country. [@Nghiem2017] also found evidence that annual GDP and unemployment rate have a significant effect on the health expenditures, however the variables used in the studies are different. Similarly, [@Factor2015] has found that country with high level of corruption have a lower spending on health care as a percentage of GDP. In a similar study, [@Giammanco2019] have found that FDI is positively related to factors related to public health. Therefore, our result somewhat follow the general trend of findings in the field of health economics. In another study, [@Wang2015] argues that the impact of health expenditures on economic growth is somewhat ambiguous.

Our study showed that factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure. Among these variables, age dependency ratio was expected to have a significant impact as higher pressure on the working population to support the non-active population reduces the chances of surplus revenues to invest in different sectors of the economy.

# Conclusions and Policy implications

The results from our study concludes that corruption GDP growth, control of corruption, and  unemployment rate have a significant and an important role in determining the amount of money spent on health sector by the country. Therefore, if we are to improve the investment on health sectors and therefore improve the citizen's access to the heath services we also need to consider the unemployment rate, corruption and the GDP growth as well. This information could be handy for donor agencies and INGO's who work to improve access of health cervice in different nations.

A special consideration should be given to nations that have an increasing GDP, high corruption, and high unemployment rate, as these nations have lower investments in health sectors. These type of countries should therefore be priotized by world bank, World Health Organization, and other institutions who allocate their resources to help countries while also considering the fact that corruption and high unemployment should also be addressed as they show causal effect.

# Limitations

Our analysis covers 2010-2020, excluding the full impact of the COVID-19 pandemic, which dramatically altered health expenditure patterns worldwide. Our OLS approach assumes linear relationships that may not capture complex interactions present in health systems, and although we employ Random Forest to address this, the "black box" nature of machine learning models limits policy interpretability. As with time-series datasets for countries, our datasets had a lot of missing values for a lot of variables across different countries. This is quite common as countries differ in their time and duration of census and surveys and not all countries folllow the same procedure to collect informations. Also, not all countries would collect the same set of information. We had to drop the missing values due to which we lost a lot of information which coould have been a valuable source of information.

# Data aquisition

The data for this research is available at the official World Bank- Data Bank website accessible [here. ](https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#) It is a free database managed by world bank group and can be accessed anytime. We can apply various filters as per our need and export the data as a .CSV or an .xlsx file.